---
title: "Legendary Pokemon"
author: "Lisa Weinstein"
date: "10/11/2019"
output: html_document
---

From the legend_by_height_plot there are more legendary pokemon with a higher weight.
```{r}
library(tidyverse)
library(dplyr)
# Import the dataset and convert variables
pokedex <- read_csv("~/Desktop/pokedex.csv", 
                    col_types = cols(name = col_factor(), 
                                     type = col_factor(),
                                     is_legendary = col_factor()))

#Examining the structure
str(pokedex)

# Prepare the data
legendary_pokemon <- pokedex %>% 
  count(is_legendary) %>% 
  mutate(prop = n / sum(n))
#printing the data frame
legendary_pokemon
# Prepare the plot
legend_by_heightweight_plot <- pokedex %>% 
  ggplot(aes(x=height_m, y=weight_kg)) +
  geom_point(aes(color=is_legendary), size = 2) +
  geom_text(aes(label = ifelse(height_m > 7.5 | weight_kg > 600, as.character(name), '')), 
            vjust = 0, hjust = 0) +
  geom_smooth(method = "lm", se = FALSE, col = "black", linetype = "dashed") +
  expand_limits(x=16) +
  labs(title = "Legendary Pokemon by height and weight",
       x = "Height (m)",
       y = "Weight (kg)") +
  guides(color = guide_legend(title = "Pokemon status")) +
  scale_color_manual(labels = c("Non-Legendary", "Legendary"),
                     values = c("#F8766D", "#00BFC4"))
# Print the plot
#labelled legendary_plot.png
legend_by_heightweight_plot
# Prepare the data
```
There are flying legendary pokemon than any other type of pokemon from looking at the barchart. 

```{r}
legend_by_type <- pokedex %>% 
  group_by(type) %>% 
  mutate(is_legendary = as.numeric(is_legendary) - 1) %>% 
  summarise(prop_legendary = mean(is_legendary)) %>% 
  ungroup() %>% 
  mutate(type = fct_reorder(type, prop_legendary))

# Prepare the plot
legend_by_type_plot <- legend_by_type %>% 
  ggplot(aes(x = type, y = prop_legendary, fill = prop_legendary)) + 
  geom_col() +
  labs(title = "Legendary Pokemon by type") +
  coord_flip() +
  guides(fill = FALSE)

# Print the plot
#labelled Legendary_Pokemon_Type_barchart.png
legend_by_type_plot
```
from looking at the boxplots, legendary pokemon are better in attack, sp_attack, defense, sp_defense, hp, speed than non-legendary pokemon. 

```{r}
# Prepare the data
legend_by_stats <- pokedex  %>% 
  select(is_legendary, attack, sp_attack, defense, sp_defense, hp, speed)  %>% 
  gather(key = "fght_stats", value = "value", -is_legendary) 

# Prepare the plot
legend_by_stats_plot <- legend_by_stats %>% 
  ggplot(aes(x = is_legendary, y = value, fill = is_legendary)) +
  geom_boxplot(varwidth = TRUE) +
  facet_wrap(~fght_stats) +
  labs(title = "Pokemon fight statistics",
       x = "Legendary status") +
  guides(fill = NULL)

# Print the plot
#labelled as boxplots.png
legend_by_stats_plot
```


```{r}
# Set seed for reproducibility
set.seed(1234)
# Save number of rows in dataset
n = nrow(pokedex)
# Generate 60% sample of rows
sample_rows <- sample(n, n*0.6)

#Create training set
pokedex_train <- pokedex  %>% 
  filter(row_number() %in% sample_rows)

# Create test set
pokedex_test <- pokedex  %>% 
  filter(!row_number() %in% sample_rows)
library(rpart)
library(rpart.plot)
# Fit decision tree
model_tree <- rpart(is_legendary ~ attack + defense + height_m + 
                    hp + sp_attack + sp_defense + speed + type + weight_kg,
                    data = pokedex_train,
                    method = "class",
                    na.action = na.omit)
rpart.plot(model_tree)

library(randomForest)
set.seed(1234)
# Fit random forest
model_forest <- randomForest(is_legendary ~ attack + defense + height_m + hp + sp_attack + sp_defense + speed + type + weight_kg,
                             data = pokedex_train,
                             importance = TRUE,
                             na.action = na.omit)

# Print model output
#labelled classification_decision_tree.png
#baseline fit to compare against the results of the random forest
print(model_forest)
```
The random forest plot has a higher true positive rate than the decision tree

```{r}
library(ROCR)

# Create prediction and performance objects for the decision tree
probs_tree <- predict(model_tree, pokedex_test, type = "prob")
pred_tree <- prediction(probs_tree[,2], pokedex_test$is_legendary)
perf_tree <- performance(pred_tree, "tpr", "fpr")

# Create prediction and performance objects for the random forest
probs_forest <- predict(model_forest, pokedex_test, type = "prob")
pred_forest <- prediction(probs_forest[,2], pokedex_test$is_legendary)
perf_forest <- performance(pred_forest, "tpr", "fpr")

# Plot the ROC curves: first for the decision tree, then for the random forest
#labelled ROC_curves.png
plot(perf_tree, col = "red", main = "ROC curves")
plot(perf_forest, add = TRUE, col = "blue")
legend(x = "bottomright",  legend = c("Decision Tree", "Random Forest"), fill = c("red", "blue"))
# Print variable importance measures
importance_forest <- importance(model_forest)
importance_forest

# Create a dotchart of variable importance
#labelled model_forest.png
varImpPlot_forest <- varImpPlot(model_forest)
```
After looking at the MeanDecreaseAccuracy plot attack is more important than defense. Also, weight_kg is more important than height_m. 

When you look at the MeanDecreaseGini plot defense is more important than attack and weight_kg is more important than height_m.
